A survey on NoSQL and NewSQL datastores

### 2014-12-06 NoSQL_table/NoSQL_table.tex
Mongo replication & storage, hbase api & replication

\multicolumn{1}{|l|}{\textbf{Distributed}}               & Master-slave, asynchronous replication   & Multi-master  & Masterless, asynchronous replication     & Master-slave or multi-master, asynchronous replication.                              & Y, not 100\% symmetric              & Masterless, updates executed on all replicas at the same time.             \\ \midrule
\multicolumn{1}{|l|}{\textbf{Storage}}                   & 16MB file limit or GridFS    & Standard Unix \& Windows FS's  & Cassandra File System (HDFS compatible)     & HDFS & HDFS or HBase & \\ \midrule

>>>>

\multicolumn{1}{|l|}{\textbf{Distributed}}               & Master-slave, asynchronous replication   & Multi-master, asynchronous replication  & Masterless, asynchronous replication     & Master-slave or multi-master, asynchronous replication.                              & Y, not 100\% symmetric              & Masterless, updates executed on all replicas at the same time.             \\ \midrule
\multicolumn{1}{|l|}{\textbf{Storage}}                   & Standard FS's, 16MB document limit    & Standard Unix \& Windows FS's  & Cassandra File System (HDFS compatible)     & HDFS & HDFS or HBase & \\ \midrule

### 2014-12-06 paper_dec/paper_dec.tex
Mongo replication & storage, hbase api & replication

Scientific progress has caused the cost of genome sequencing to drop at an exponential rate over the past decade and a half, even outpacing Moore's law since 2008 \cite{wetterstrand_sequencing_cost}. As in all sorts of biological, medical and pharmaceutical research, more and more genomes are being sequenced the amount of data generated increases rapidly. For instance, the whole genome sequencing pipeline of the Broad Institute\cite{broad_institute}, a reference in the field, generates in the order of 3 TB of intermediary data when sequencing a single human genome. Also, when sequenced with 50x mean coverage (denoting the average number of times a base has been read in the sequencing process\cite{coverage_definition}) a single human genome takes 50 GB to store in compressed format,
%TODO source
so as this scales to genomes of millions of people and other organisms, storage and processing requirements become ever more demanding in terms of scalability, latency and concurrency.
\\A logical evolution has been to tackle these problems with high performance computing systems. The Exascience Life Lab of imec, Intel, Janssen Pharmaceutica and 5 Flemish universities, actively researches the application of supercomputers for processing whole genome sequences \cite{exascience_life_lab}\cite{lifelab_BWA}.
\\Increasing popularity of web services such as social networks has caused a similar explosion of data for web companies. To handle this so-called Big Data\cite{mashey1997big} in an adequate way, traditional relational DBMS no longer suffice. Therefore, large web companies such as Google and Amazon have developped new storage solutions that meet the demands for high and incremental scalability, low latency and high availability\cite{baker2011megastore}. This has spawned many so called NoSQL ('Not only SQL') databases, which loosen the rigid relational datamodel in favour of better scaling and easier distribution of the data.
\\NoSQL datastores come in multiple flavours and can be divided in a few categories based on the datamodel they use:
\begin{itemize}
\item Columnar stores: based on the datamodel pioneered by Google's BigTable, these store data in "a sparse, distributed, persistent multidimensional sorted map" \cite{chang2008bigtable}. In BigTable's case this is a map of row key, column key and a timestamp. In this way, multiple versions of the data can be stored in chronological order. Because the system doesn't interpret the data, relationships can't be modeled. This is left to application logic\cite{hecht2011nosql}.
\item Document stores: these store data as key-value pairs encapsulated in documents. Values can be of a wide variety of types, such as nested documents, lists or scalar values. Attribute names can be dynamically specified at runtime and need not adhere to a fixed schema\cite{cattell2011scalable}. This is well suited for modelling complex data structures. Many document stores use the JSON-fileformat (or some derived form). In contrast to columnar stores, the values in documents are not opaque to the system and can thus be queried and indexed\cite{hecht2011nosql}.\\
The term NewSQL data stores is being used to classify a set of solutions that aim to combine the scalability, distribution and fault tolerance of NoSQL stores with the relational data model. Though they all use the relational model and offer SQL, NewSQL stores vary greatly under the hood, depending on the architecture they are built on\cite{grolinger2013data}. This paper proceeds first by elaborating on the methodology used to select and compare the chosen databases and then by reviewing each of the six datastores in detail.
\end{itemize}

@@

%\include{tabel}
\section{Document stores}

@@

MongoDB stores data in BSON (binary JSON) documents. It has powerful indexing support, with the ability to define secondary indexes of a wide array of types on all attributes, much like in the traditional relational model. These are implemented using B-trees \cite{mongodb_indexes}. To aid with denormalization, the documents can contain embedded documents and arrays, thus obviating the need for joins in the query language.
\\MongoDB offers API's in many languages and the functionality to define the equivalent of SQL WHERE-clauses as javascript expressions. These are then translated to MongoDB's proprietary internal querying language \cite{grolinger2013data}. The MongoDB query optimizer processes queries and chooses the most efficient plan for a query given the available indexes. These plans are cached if there are multiple viable options and can be reevaluated as the data evolves \cite{mongodb_query_plans}.
\\The consistency of MongoDB is configurable, strong consistency can be attained in two ways: setting the connection to read-only from the master node (which has the most up-to-date version of the data), or forcing a write to succeed only after all replicas have acknowledged it. The former degrades the scaling ability of read requests, the latter the latency of write requests \cite{grolinger2013data}.
For concurrency control, MongoDB offers single-document atomicity and implements reader-writer locks. Having to lock on writes severely impacts performance in write-intensive scenarios.
\\\\As a wrap-up, MongoDB stores BSON-files accessible through many API's with flexible querying and indexing techniques, but its concurrency and consistency model have some drawbacks.

@@

Apache HBase is an open source datastore using Google BigTable's datamodel, but running on top of the Hadoop Distributed File System (HDFS) instead of the Google File System (GFS).
\\Since its launch, HBase has adopted several secondary indexing mechanisms. For indexing, HBase also uses LSM trees\cite{borthakur2011apache}\cite{sears2012blsm}, and it also provides Bloom filters\cite{hbase_schema}. HBase doesn't offer an SQL-like advanced querying language, but because of its HDFS underpinnings it can function as both input and output for MapReduce jobs.
\\The strong points of HBase are its strong consistency, which is rare among NoSQL-stores, and its concurrency model: ACID-compliant single row transactions and optimistic multi-version concurrency control for wider scope operations\cite{hbase_acid}\cite{grolinger2013data}\cite{borthakur2011apache}.\\\\
In conclusion, HBase offers flexible data-modeling and is especially useful when there is already a HDFS dataset and when MapReduce compatibility is desired. It scales well to very large datasets and offers excellent concurrency control and strong consistency.

>>>>

Scientific progress has caused the cost of genome sequencing to drop at an exponential rate over the past decade and a half, even outpacing Moore's law since 2008 \cite{wetterstrand_sequencing_cost}. Because in all sorts of biological, medical and pharmaceutical research, more and more genomes are being sequenced, the amount of generated data increases rapidly. For instance, the whole genome sequencing pipeline of the Broad Institute\cite{broad_institute}, a reference in the field, generates in the order of 3 TB of intermediary data when sequencing a single human genome. Also, when sequenced with 50x mean coverage (denoting the average number of times a base has been read during the process\cite{coverage_definition}) a single human genome takes 50 GB to store in compressed format,
%TODO source
so as this scales to genomes of millions of people and other organisms, storage and processing requirements become ever more demanding in terms of scalability, latency and concurrency.
\\A logical evolution has been to tackle these problems with high performance computing systems. The Exascience Life Lab of imec, Intel, Janssen Pharmaceutica and 5 Flemish universities, actively researches the application of supercomputers for processing whole genome sequences \cite{exascience_life_lab}\cite{lifelab_BWA}.
\\Increasing popularity of web services such as social networks has caused a similar explosion of data for web companies. To handle this so-called Big Data\cite{mashey1997big} in an adequate way, traditional relational DBMS no longer suffice. Therefore, large web companies such as Google and Amazon have developped new storage solutions that meet the demands for high and incremental scalability, low latency and high availability\cite{baker2011megastore}. This has spawned many so called NoSQL ('Not only SQL') databases, which loosen the rigid relational datamodel in favour of better scaling and easier distribution of the data.
\\NoSQL datastores come in multiple flavours and can be divided in a few categories based on the datamodel they use:
\begin{itemize}
\item Columnar stores: based on the datamodel pioneered by Google's BigTable, these store data in "a sparse, distributed, persistent multidimensional sorted map" \cite{chang2008bigtable}. In BigTable's case this is a map of row key, column key and a timestamp. In this way, multiple versions of the data can be stored in chronological order. Because the system doesn't interpret the data, relationships can't be modeled. This is left to application logic\cite{hecht2011nosql}.
\item Document stores: these store data as key-value pairs encapsulated in documents. Values can be of a wide variety of types, such as nested documents, lists or scalar values. Attribute names can be dynamically specified at runtime and need not adhere to a fixed schema\cite{cattell2011scalable}. This is well suited for modelling complex data structures. Many document stores use the JSON-fileformat (or some derived form). In contrast to columnar stores, the values in documents are not opaque to the system and can thus be queried and indexed\cite{hecht2011nosql}.\\
The term NewSQL data stores is being used to classify a set of solutions that aim to combine the scalability, distribution and fault tolerance of NoSQL stores with the relational data model. Though they all use the relational model and offer SQL, NewSQL stores vary greatly under the hood, depending on the architecture they are built on\cite{grolinger2013data}. This paper proceeds first by elaborating on the methodology used to select and compare the chosen databases and then by reviewing each of the six datastores in detail.
\end{itemize}

@@

\section{Document stores}

@@

MongoDB stores data in BSON (binary JSON) documents. It has powerful indexing support, with the ability to define secondary indexes of a wide array of types on all attributes, much like in the traditional relational model. These are implemented using B-trees \cite{mongodb_indexes}. To aid with denormalization, the documents can contain embedded documents and arrays, thus obviating the need for joins in the query language.
\\The document size is limited to 16 MB, to help ensure that a single document cannot take up excessive amounts of RAM or bandwidth. To store and retrieve larger files, the built-in tool GridFS (which is not an actual file system) can split up files in smaller chunks and store these chunks as separate documents\cite{mongodb_gridfs}.
\\MongoDB offers API's in many languages and the functionality to define the equivalent of SQL WHERE-clauses as javascript expressions. These are then translated to MongoDB's proprietary internal querying language \cite{grolinger2013data}. The MongoDB query optimizer processes queries and chooses the most efficient plan for a query given the available indexes. These plans are cached if there are multiple viable options and can be reevaluated as the data evolves \cite{mongodb_query_plans}.
\\The consistency of MongoDB is configurable. Strong consistency can be attained in two ways: setting the connection to read-only from the master node (which has the most up-to-date version of the data), or forcing a write to succeed only after all replicas have acknowledged it. The former degrades the scaling ability of read requests, the latter the latency of write requests \cite{grolinger2013data}.
\\Data is replicated asynchronously using range-partitioning: nodes are responsible for ranges of keys. This speeds up range queries, but can create hotspots and load-balancing issues. To route updates to the right replicas, MongoDB operates in a master-slave setting.\\
For concurrency control, MongoDB offers single-document atomicity and implements reader-writer locks. Having to lock on writes severely impacts performance in write-intensive scenarios.
\\\\As a wrap-up, MongoDB stores BSON-files accessible through many API's with flexible querying and indexing techniques, but its concurrency and consistency model have some drawbacks.

@@

Apache HBase is an open source datastore using Google BigTable's datamodel, but running on top of the Hadoop Distributed File System (HDFS) instead of the Google File System (GFS).
\\Since its launch, HBase has adopted several secondary indexing mechanisms. For indexing, HBase also uses LSM trees\cite{borthakur2011apache}\cite{sears2012blsm}, and it also provides Bloom filters\cite{hbase_schema}. HBase comes with a Java API but without an SQL-like advanced querying language, though a workaround via Apache Hive, another data warehousing and querying project\cite{apache_hive}, and its query language HiveQL is possible. Because of its HDFS underpinnings it can function as both input and output for MapReduce jobs.
\\HBase partitions data in ranges like BigTable and replicates updates in master-slave or multi-master fashion. Read-requests are not distributed however, as rows are only serviced by one server. The replicas are solely intended for failure recovery.
\\The strong points of HBase are its strong consistency, which is rare among NoSQL-stores, and its concurrency model: ACID-compliant single row transactions and optimistic multi-version concurrency control for wider scope operations\cite{hbase_acid}\cite{grolinger2013data}\cite{borthakur2011apache}.\\\\
In conclusion, HBase offers flexible data-modeling and is especially useful when there is already a HDFS dataset and when MapReduce compatibility is desired. It scales well to very large datasets and offers excellent concurrency control and strong consistency.
